### 2025-09-18 paper/sec/authoring-workflow.tex
Tweak.
Roll out 'agent' macros.

\begin{itemize}
    \item \textbf{SuggestionAgent}. LLM-based agent which identifies text fragments potentially computable from data.
    \item \textbf{InterpretationAgent}. LLM-based agent which, for a given text fragment (provided by
    SuggestionAgent or by the author), attempts to synthesise a Fluid expression computing the target text.
\end{itemize}

@@

Figure~\ref{fig:architecture} shows the overall architecture of the system.
The author initially provides the text and accompanying data.
The SuggestionAgent analyses the input and identifies candidate fragments of natural language which could be replaced by Fluid expressions.
The user highlightes a fragment of interest, which is then sent to the InterpretationAgent, which generates a corresponding candidate expression.
Internally the InterpretationAgent uses a compiler-in-the-loop validation process to ensure the expression is well-formed.
The user is then able to accept the expression (in which case it is incorporated into the program) or reject and leave the text uninterpreted.

@@

Expressions generated by the \InterpretationAgent are evaluated by the Fluid interpreter. If evaluation fails,
the error is used to extend the prompt and the extended prompt is sent back to the LLM for regeneration. This
loopback mechanism enables iterative refinement of expressions, improving accuracy and robustness. There are
two kinds of loopback error:

\begin{itemize}
    \item \textbf{Invalid Program}. Evaluation fails due a syntax error, undeclared identifier, division by
    zero, array index out of bounds or other runtime problem.
%    \item \textbf{Invalid Expression Type}. The value of the expression is not a string.
    \item \textbf{Mismatching String}. The string value of the expression is not equal to the target text.
\end{itemize}

>>>>

\begin{itemize}
    \item \textbf{\SuggestionAgent}. LLM-based agent which identifies text fragments potentially computable from data.
    \item \textbf{\InterpretationAgent}. LLM-based agent which, for a given text fragment (provided by
    SuggestionAgent or by the author), attempts to synthesise a Fluid expression computing the target text.
\end{itemize}

@@

Figure~\ref{fig:architecture} shows the overall architecture of the system. The author initially provides the
text and accompanying data. The \SuggestionAgent analyses the input and identifies candidate fragments of
natural language which could be replaced by Fluid expressions. The user highlightes a fragment of interest,
which is then sent to the \InterpretationAgent, which generates a corresponding candidate expression.
Internally the \InterpretationAgent uses a compiler-in-the-loop validation process to ensure the expression is
well-formed. The user is then able to accept the expression (in which case it is incorporated into the
program) or reject and leave the text uninterpreted.

@@

Expressions generated by the \InterpretationAgent are evaluated by the Fluid interpreter. If evaluation fails,
the error is used to extend the prompt and the extended prompt is sent back to the LLM for regeneration. This
loopback mechanism enables iterative refinement of expressions, improving accuracy and robustness. There are
three kinds of loopback error:

\begin{itemize}
    \item \textbf{Invalid Program}. Evaluation fails due a syntax error, undeclared identifier, division by
    zero, array index out of bounds or other runtime problem.
    \item \textbf{Invalid Expression Type}. The value of the expression is not convertible to a string.
    \item \textbf{Mismatching String}. The string value of the expression is not equal to the target text.
\end{itemize}

### 2025-09-18 paper/sec/evaluation.tex
Roll out 'agent' macros.

\paragraph{Initial validation}
In the initial phase, the generated expression is executed via Fluid CLI.
If the execution is successful, the result obtained is compared with the expected result.
In case of error, the output is forwarded to the interpretation agent, which proceeds to regenerate the expression.

\paragraph{Validation via counterfactual testing}
In the second phase, the data used for execution is deliberately modified in order to verify the consistency of the behaviours.
Both the expected and generated expressions are executed again and the results are compared.
If discrepancies emerge, the error is forwarded to the interpretation agent.
If there are no differences, the expression is considered valid and is returned to the user.

>>>>

\paragraph{Initial validation}
In the initial phase, the generated expression is executed via Fluid CLI. If the execution is successful, the
result obtained is compared with the expected result. In case of error, the output is forwarded to the
\InterpretationAgent, which proceeds to regenerate the expression.

\paragraph{Validation via counterfactual testing}
In the second phase, the data used for execution is deliberately modified in order to verify the consistency
of the behaviours. Both the expected and generated expressions are executed again and the results are
compared. If discrepancies emerge, the error is forwarded to the \InterpretationAgent. If there are no
differences, the expression is considered valid and is returned to the user.
